MSB104_assignment1

### 2023-10-10 MSB104_assignment1.qmd
Final commit 10.10.23

#### Description of the numbers from German sub-regions

As mentioned earlier, the data include numbers from NUTS3 level

Som beskrevet over inneholder tallene fra NUTS3 level disse dataene.
Vi bestemte oss for å dele Tyskland inn i Nord og Sør pga.
(se comments).

Gå gjennom outliers ved å bruke Espen sitt tilsendte dokument.
F.eks.
beskrive Wolfsbrug, Hamburg etc.
Beskriv de laveste, si noe om hvorfor dette er tilfelle.

Beskrive finanskrisen og Covid i grafen.

Beskrive deskriptiv statistikk.

@@

@@

>>>>

#### Description of the numbers from German regions

In order to achieve a presentable figure showing GDP per year for the German sub-regions, we decided to divide the data into Northern and Southern Germany.
This prevented the data from becoming overly clustered and improved our ability to interpret the graphs.
The first figure presents data for North Germany, and the second figure presents data for South Germany.

In year 2000, the region with highest GDP per capita was Hamburg (DE60), which is located in North Germany, with a GDP of 45,659 €.
Hamburg's high GDP in Germany can be attributed to several key factors.
First, Hamburg is home to a diverse range of industries, including maritime trade, logistics, manufacturing, aviation, and media.
This economic diversification contributes to a robust and stable economy.
Second, the Port of Hamburg is one of Europe's largest and busiest ports.
It serves as a major gateway for international trade, making Hamburg a crucial hub for import and export activities @imfnd.

The region with the lowest GDP per capita was Thüringen (DEG0), which is located in North Germany, with a GDP of 16,139 €.
Thüringen has one of the lowest GDPs in Germany due to several factors.
First, Thüringen has a relatively small and less diversified economy compared to other German states.
It relies heavily on traditional industries such as agriculture and manufacturing, which may not contribute significantly to GDP.
Second, Thüringen has a lower population compared to other regions.
A smaller population can lead to reduced economic output @thuringennd.

In Germany in 2000, GDP per capita had an average value (mean) of 25,593 €.
The median value of 24,894 € indicates that the distribution is slightly skewed to the left, with about half of the population having values below the mean.
The standard deviation of 5,872 € quantifies the extent of variation around the mean, suggesting a moderate level of variability in GDP per capita across the population.

In year 2020, the region with highest GDP per capita was still Hamburg (DE60), with a GDP of 64,366 €, and the region with the lowest GDP per capita was still Thüringen (DEG0), with a GDP of 28,874 €.

In Germany in 2020, GDP per capita had an average value (mean) of 39,439 €.
The median value of 38,941 € indicates that the distribution is slightly skewed to the left, with about half of the population having values below the mean.
The higher standard deviation of 8,267 € compared to 2000 quantifies the extent of variation around the mean, suggesting a relatively higher level of variability in economic indicators across the population in 2020.

We can also observe that there is a slope in both figures from 2007 to 2009.
This is most likely caused by the Financial Crisis (2007-2008), which caused the GDP per capita to decline.
There is also a slope in both figures in 2019, which is most likely due to Covid-19.

@@

#### Description of the numbers from Finnish regions

In year 2000, the region with highest GDP per capita was Helsinki-Uusimaa (FI1B), with a GDP of 36,478 €.
Helsinki-Uusimaa is one of the regions in Finland with the highest GDP per capita for several reasons.
Helsinki, the capital of Finland, is located in the Uusimaa region.
Being the largest city and the national capital, Helsinki serves as the economic, financial, and cultural center of Finland.
It is home to numerous businesses, corporate headquarters, government institutions, and educational and research facilities.
The presence of these institutions contributes significantly to the region's economic activity (**SOURCE**).

The region with the lowest GDP per capita was Pohjois- ja Itä-Suomi (FI1D), with a GDP of 20,605 €.
Pohjois- ja Itä-Suomi is one of the regions in Finland with lower GDP per capita for several reasons.
For instance, its located in Northern and Eastern Finland, which characterized by their vast, sparsely populated areas and harsh climatic conditions.
These factors can limit economic activities in certain sectors, such as agriculture and construction, which may thrive more in milder climates and more densely populated regions (**SOURCE**).

In Finland in 2000, while the average GDP per capita was 27,201 €, there was moderate income inequality, with some segments of the population earning significantly more or less than the average.
The difference between the mean (€27,201) and median (24,523 €) implies the presence of relatively high-income individuals or regions that influenced the mean value.

In year 2020, the region with highest GDP per capita was still Helsinki-Uusimaa (FI1B), with a GDP of 55,750 €, and the region with the lowest GDP per capita was still Pohjois- ja Itä-Suomi (FI1D), with a GDP of 36,072 €.

In Finland in 2020, while the average GDP per capita was 41,669 €, there was a noticeable level of income inequality.
Some segments of the population earned significantly more or less than the average, as evidenced by the difference between the mean (41,669 €) and median (38,178€).
The presence of a right-skewed distribution indicates the influence of relatively high-income individuals or regions on the mean value.

Also in the graphs for Finland we see the effects on GDP of the Financial Crisis (2007-2008) and Covid-19.

@@

#### Description of the numbers from Romanian regions

In year 2000, the region with highest GDP per capita was București-Ilfov (RO32), with a GDP of 3972 €.
București-Ilfov is the region with the highest GDP in Romania for several reasons.
București (Bucharest) is not only the capital city of Romania but also its economic and financial center.
As the capital, it serves as the hub for government institutions, businesses, and financial organizations.
This concentration of administrative and economic activity contributes significantly to the region's GDP (**SOURCE**).

The region with the lowest GDP per capita was Nord-Est (RO21), with a GDP of 1253 €.
Nord-Est is one of the regions in Romania with the lowest GDP for several reasons.
Nord-Est has an economy that is predominantly based on agriculture, which tends to have lower productivity and lower value-added compared to other sectors such as manufacturing, services, or technology.
This agricultural dependency can limit the region's overall economic output (**SOURCE**).

In Romania in 2000, the average GDP per capita was 1889 €, however there was significant variation in income levels across different regions or segments of the population.
The median (1619 €), being lower than the mean (1889 €), hints at the presence of income inequality or disparities in economic well-being.

In year 2020, the region with highest GDP per capita was still București-Ilfov (RO32), with a GDP of 26,270 €, and the region with the lowest GDP per capita was still Nord-Est (RO21), with a GDP of 7478 €.

In Romania in 2020, the average GDP per capita was around 11,628 €, however there is still significant variation in income levels across different regions or segments of the population.
The median (9822 €), being lower than the mean (11,628 €), hints at the presence of income inequality or disparities in economic well-being.

Also in the graphs for Romania we see the effects on GDP of the Financial Crisis (2007-2008) and Covid-19.
